ProbabilityBoundsAnalysis.jl:\\ Arithmetic with sets of distributions

### 2022-04-14 paper/paper.tex
further removes footnotes

The goal of \texttt{ProbabilityBoundsAnalysis.jl}\footnote{https://github.com/AnderGray/ProbabilityBoundsAnalysis.jl} is to \textbf{compute guaranteed bounds on functions of random variables given only partial knowledge of the input probability distributions and their dependencies}. That is, to compute with partial knowledge about the input joint distribution. Ideally, all of the available information about random variables should be used, but without making any extra assumptions beyond what are justified by what is known.

@@

  \caption{A precise distribution, an interval and a probability box}
  \label{fig:figure1}

@@

  \caption{Various outer representations of distribution functions, left $4$ steps and centre $25$ steps, and a p-box (right) with 10 steps}
  \label{fig:figure2}

@@

\fi
\section{Probability Boxes}
\label{sec:pboxes}
A real-valued random variable is characterised by its distribution function $F$, which is a monotonically increasing function from the real numbers onto the interval $[0, 1]$ such that the value of the function at negative infinity is zero and the value of the function at positive infinity is one.  A p-box consists of a pair of such functions that are used to circumscribe an imprecisely known distribution function $F$. The p-box, in its simplest form consisting of the pair of bounding functions, identifies a set of probability distributions that lie entirely within these bounds. Additional information about the random variable may be available, such as bounds on its mean and variance and its distribution family, which may be used to further restrict the set of distributions. A p-box is thus defined by the following three constraints: (1) two bounding cumulative distribution functions (cdf), (2) interval bounds on the mean and variance, and (3) a collection of distribution families:
\begin{enumerate}
  \item $\begin{aligned}[t]
    \underline{F}(x) \leq F(x) \leq \overline{F}(x), \\%&&&& \text{(Bounds on the cdf)} \\
  \end{aligned}$
  \item $\begin{aligned}[t]
    \mu &\in [\underline{ \mu }, \overline{ \mu }]  \\%&&&& \text{(bounds on mean and variance)}\\
    \sigma^2 &\in [\underline{\sigma}^2 , \overline{\sigma}^2],
  \end{aligned}$
  \item $\begin{aligned}[t]
      F \in \bm{F}. \\%&&&&&&& \text{(distribution family)} \\
  \end{aligned}$
  \end{enumerate}
\noindent That is, a random variable is a member of a p-box if its cdf $F$ falls within the cdf bounds of the p-box $F(x) \in [\underline{F}(x), \overline{F}(x)]$ for all $x$, its moments are inside the interval moments of the p-box, and it belongs to a family of distribution functions (e.g. normal, uniform) considered by the p-box. Some of the constraints may be missing. For example, if the distribution family is unknown then the set is defined solely from the cdf and moment bounds. Such p-boxes are sometimes called non-parametric, since its members do not belong to any particular class of distribution. Some constraints may also be inferred from others. For example, the interval moments may be bounded from the cdf bounds, and cdf may be bounded from moment information (explored further in section \ref{section:Moments}). Figure\footnote{Scripts for all figures can be found in /examples/JuliaCon.} \ref{fig:figure1} shows an example of a distribution function, interval and a p-box. The grey shaded region in the interval and p-box bounds an infinite collection of probability distributions.

@@

  \caption{Various parametrically constructed p-boxes with interval parameters}
  \label{fig:figure3}
\end{figure*}
P-boxes generalise precise distributions and intervals in the following way. A distribution is a p-box with a precise cdf and moments, i.e. when
\begin{align*}
  \underline{F}(x) &= \overline{F}(x), \\ 
  \underline{\mu}  &= \overline{\mu}, \\ 
  \underline{\sigma}^2 &= \overline{\sigma}^2 .
\end{align*}
\noindent An interval $[a,b]$ is a p-box whose bounds are step functions
\begin{align*}
    \underline{F}(x) &= \epsilon_{b}(x) ,\\
    \overline{F}(x) &= \epsilon_{a}(x) ,
\end{align*}
\noindent where $\epsilon_k$ is
\begin{equation*}
   \epsilon_k(x) = \begin{cases} 0 &\text{when } x < k \\ 1 &\text{when } x \geq k. \end{cases} 
\end{equation*}
\noindent Moreover, theoretical bounds on the mean and variance of an interval can be found \cite{ferson2002ramas}
\begin{align*}
  \mu &\in [a, b], \\
  \sigma^2 &\in [0, (b - a)^{2}/4].
\end{align*}
\noindent That is, it is not possible to find a distribution inside $[a, b]$ with a variance greater than $(b-a)^{2}/4$. The lower bound on the variance is zero, since scalars are also included in the interval. Under this definition of an interval, a random sample will always be the interval $[a, b]$, the cdf returns 0 when $x < a$, the interval $[0,1]$ when $a \leq x < b$, and $1$ when $b \leq x$. Further, the probability measure of the interval $X = [a, b]$ will be: 
\begin{equation*}
  \mathbb{P}_{X}(U) = \begin{cases}
    0 & \text{when } U \cap [a,b] = \emptyset \\
    1 & \text{when } [a,b] \subseteq U  \\
    [0, 1] & \text{otherwise.}\\
  \end{cases}
\end{equation*}
\noindent That is, if the set $U$ does not intersect the interval $X$, $\mathbb{P}_{X}=0$, i.e. $U$ certainly does not contain any of the random variables. If $X$ is fully contained in $U$, $\mathbb{P}_{X}=1$, i.e. $U$ certainly contains all of the random variables. And finally if $U$ intersects (but does not contain) $X$, the probability measure is the vacuous probability interval $\mathbb{P}_{X} = [0, 1]$, i.e. we are completely uncertain about the containment. Note that the $P_{X} = \{0, 1, [0,1]\}$ is due to the interval bounds being step functions. Generally p-boxes may yield other probability intervals.
\subsection{Outer representations of p-boxes}
\label{sec:outer_approx}
An important feature of probability bounds analysis is how distribution functions and p-boxes are represented. Generally, analytical solutions for cdfs are not readily available, for example the normal distribution's cdf can only be found by integrating the probability density, usually with quadrature. However, even if the functions were available analytically, when the variables are used in an arithmetic operation the output distribution will not necessarily belong to the same family. Therefore, we require a representation of these continuous functions which is not dependent on the distribution family (for example as used in polynomial chaos expansion), and ideally is robust in the sense that it can produce an interval error for the numerical representation. Such a representation of a probability is sometimes called an "outer approximation", in the same fashion an interval is an outer approximation of a floating point number. Such outer approximations contract to the exact value or function as more computational resources are used.

@@

A minimal\footnote{Code can be found in examples/minimal.} p-box structure is as follows
\begin{lstlisting}[language = Julia]
  using IntervalArithmetic

@@

    return pbox(u_z_, d_z_, Z_moms.mean, 
    Z_moms.var, F.shape)
  end
\end{lstlisting}
\noindent How moment transformations are performed are not described here, however they are available in \texttt{MomentArithmetic.jl}\footnote{https://github.com/AnderGray/MomentArithmetic.jl} \cite{ferson2021distribution}. Figure \ref{fig:figure4} shows an example of an operation between a precise distribution and an interval.
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_dist.pdf}
  \raisebox{8.0mm}{\noindent\Large*}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_in.pdf}
  \raisebox{9.0mm}{{\Large$\rightarrow$}}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_pbox.pdf}
  \caption{Illustration of a p-box resulting from a binary operation between a distribution and an interval.}
  \label{fig:figure4}
\end{figure}
\subsubsection{When dependencies between distributions are unknown}
In the case where a binary operation is performed between two distributions but whose dependence or correlation is unknown, a single output distribution cannot be defined. However a bounding p-box can be computed. This is the original question of Kolmogorov as described in the introduction. We call such operations between p-boxes where the marginals are fixed but no joint information is known \textit{Fréchet} operations, after the mathematician Maurice R. Fréchet, who first discovered how to bound joint distributions with fixed marginals \cite{frechet1935generalisation}. Figure \ref{fig:figure5} illustrates a Fréchet operation between a uniform and a normal distribution.
\begin{figure}[htp]

@@

A bivariate copula (2-copula) $C$ is any function $C:[0,1]^2 \rightarrow [0,  1]$ with the following properties
\begin{enumerate}
  \item Grounded: $C(0,v) = C(u,0) = 0$,
  \item Uniform margins: $C(u,1) = u; \;C(1,v) = v$,
  \item 2-increasing: \\ $C(u_{2},v_{2}) - C_{}(u_{2}, v_{1}) - C(u_{1}, v_{2}) + C(u_{1}, v_{1}) \ge 0$\\ for all $0 \le u_{1} \le u_{2} \le 1$ and $0 \le v_{1} \le v_{2} \le 1$ .
\end{enumerate}
\noindent Three important 2-copulas are
\begin{align*}
  W(u,v) &= \mathrm{max}( u + v-1,0),  \\
  \Pi(u,v) &= uv, \\
  M(u,v) &= \mathrm{min}(u,v),
\end{align*}
\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.45\textwidth]{../examples/JuliaCon/fig7/biv_cdf.pdf}\hfill
  \includegraphics[width=0.45\textwidth]{../examples/JuliaCon/fig7/samples.pdf}
  \caption{Cdf bounds of a bivariate p-box (top) with $X \sim \text{Beta}([4,8], 3)$, $Y \sim \text{N}([0,2], 2)$, and $\rho_{XY} = -0.8$, and 100 samples (bottom)}
  \label{fig:figure7}
\end{figure*}
\noindent where $W$ encodes maximally negative correlation, $\Pi$ encodes stochastic independence, and $M$ encodes maximally positive correlation. Moreover, $W$ and $M$ are bounds on all 2-copulas: $W(u,v) \leq C(u,v) \leq M(u,v)$. A bivariate cdf $F_{XY}$ can be constructed from two marginal cdfs $F_{X}$ and $F_{Y}$ and copula $C$ using Sklar's theorem
\begin{equation*}
  F_{XY}(x,y) = C(F_{X}(x), F_{Y}(y)),
\end{equation*}
\noindent where the constructed bivariate distribution will follow the dependence encoded in the copula and have the prescribed marginals. Sklar's theorem also works in reverse, where the dependence can be extracted from a multivariate distribution: 
\begin{equation*}
  C(u,v) = F_{XY}(F^{-1}_{X}(u), F^{-1}_{Y}(v)).
\end{equation*}
\noindent Copulas also give a very convenient way to sample from $F_{XY}$ in a generalisation of inverse transform sampling, where first the copula is sampled $(u,v) \sim C$, and then transformed through the inverses of the marginals $(F^{-1}_{X}(u), F^{-1}_{Y}(v))$.

@@

  \caption{Top left shows an initial p-box, and others show results of unary operations on this p-box}
  \label{fig:figure8}

\end{figure*}
A minimal structure, cdf, random sample and probability measure of a bivariate p-box is \footnote{https://github.com/AnderGray/BivariateCopulas.jl}
\begin{lstlisting}[language = Julia]
  using BivariateCopulas
  using BivariateCopulas: sample

@@

  \caption{Top left shows two initial p-boxes $X, Y \sim U([1,2], 4)$, and others show results of binary operations on these two p-boxes with different dependencies. Each colour represents a different copula. The colour code is $C = \{$ {\color{purple} $W$} (opposite), {\color{green} $\Phi_{-0.5}$ }(correlation -0.5), {\color{red} $\Pi$} (independence), {\color{orange} $\Phi_{0.5}$} (correlation 0.5), {\color{blue} $M$}(perfect) $\}$}
  \label{fig:figure9}
\end{figure*}
A binary operation between two random variables $X$ and $Y$ with cdfs $F_{X}$ and $F_{Y}$ and copula $C$ can be computed with the following Lebesgue-Stieltjes integral
\begin{equation*}
  \sigma_{C,L}(z) = \int_{L\{z\}}dC(F_{X}(x),F_{Y}(y)) ,
\end{equation*}
where the integration domain is the set $L\{z\} = \{(x,y)| x,y \in \mathbb{R}, L(x,y) < z\}$ (the set of all $x$ and $y$ for which $L(x,y) < z$), and where $L$ is some binary operation (e.g. $L(x,y) = x+y$). The above $\sigma_{C,L}$ convolution is written for precise distributions, however can be extended for p-boxes by performing the convolution with some combination of the p-boxes bounds in the style of interval arithmetic, e.g.,
\begin{eqnarray*}
  + 
  \begin{cases}
    \underline{F_Z} = \underline{F_X} + \underline{F_Y} \vspace{2mm}\\
    \overline{F_Z} = \overline{F_X} + \overline{F_Y}
  \end{cases},\;
  -\begin{cases}
    \underline{F_Z} = \underline{F_X} - \overline{F_Y}\vspace{2mm}\\
    \overline{F_Z} = \overline{F_X} - \underline{F_Y}
  \end{cases} .
\end{eqnarray*}
Williamson and Downs provide an algorithm to compute rigorous bounds on the above convolution, but only under independence (when $C = \Pi$). Around the same time, Yager \cite{yager1986arithmetic} described a similar style convolution on two belief functions under the assumption independence, which is very similar to that of Williamson and Downs'. We therefore provide a generalised version of Yager's algorithm where any copula $C$ can be specified. Similar to a unary operation, a binary operation between two p-boxes can be performed by first transforming the bivariate p-box into a two dimensional belief function (which has interval boxes as focal elements), evaluating the operation on the focal elements with interval arithmetic, and transforming the resulting belief function (which will be univariate after the operation) back to a p-box. Converting a bivariate p-box into a 2D belief function is similar to that of the 1D case, however this time the mass assignment $m$ will be determined by the copula $C$. For two univariate focal elements $X$ and $Y$ with probability widths $p_{x} = [p_{x1}, p_{x2}]$ and $p_{y} = [p_{y1}, p_{y2}]$ (i.e. their vertical width in Figure \ref{fig:figure2}, with their 1D mass assignment being $p_{x2} - p_{x1}$), the 2D mass assignment of the interval box $X \times Y$ will be
\begin{equation*}
  C(p_{x2}, p_{y2}) - C(p_{x1}, p_{y2}) - C(p_{x2}, p_{y1}) + C(p_{x1}, p_{y1}) .
\end{equation*}
The following function converts a bivariate p-box into a 2D belief function
\begin{lstlisting}[language = Julia]
  function split(F :: BivPbox)
    X = F.X; Y = F.Y; C = F.C

>>>>

The goal of \texttt{ProbabilityBoundsAnalysis.jl} is to \textbf{compute guaranteed bounds on functions of random variables given only partial knowledge of the input probability distributions and their dependencies}. That is, to compute with partial knowledge about the input joint distribution. Ideally, all of the available information about random variables should be used, but without making any extra assumptions beyond what are justified by what is known.

@@

  \caption{A precise distribution, interval and probability box.}
  \label{fig:figure1}

@@

  \caption{Various outer representations of distribution functions, left $4$ steps and centre $25$ steps, and a p-box (right) with 10 steps.}
  \label{fig:figure2}

@@

\fi
\section{Probability Boxes}
\label{sec:pboxes}
A real-valued random variable is characterised by its distribution function $F$, which is a monotonically increasing function from the real numbers onto the interval $[0, 1]$ such that the value of the function at negative infinity is zero and the value of the function at positive infinity is one.  A p-box consists of a pair of such functions that are used to circumscribe an imprecisely known distribution function $F$. The p-box, in its simplest form consisting of the pair of bounding functions, identifies a set of probability distributions that lie entirely within these bounds. Additional information about the random variable may be available, such as bounds on its mean and variance and its distribution family, which may be used to further restrict the set of distributions. A p-box is thus defined by the following three constraints: (1) two bounding cumulative distribution functions (cdf), (2) interval bounds on the mean and variance, and (3) a collection of distribution families:
\begin{enumerate}
  \item $\begin{aligned}[t]
    \underline{F}(x) \leq F(x) \leq \overline{F}(x), \\%&&&& \text{(Bounds on the cdf)} \\
  \end{aligned}$
  \item $\begin{aligned}[t]
    \mu &\in [\underline{ \mu }, \overline{ \mu }]  \\%&&&& \text{(bounds on mean and variance)}\\
    \sigma^2 &\in [\underline{\sigma}^2 , \overline{\sigma}^2],
  \end{aligned}$
  \item $\begin{aligned}[t]
      F \in \bm{F}. \\%&&&&&&& \text{(distribution family)} \\
  \end{aligned}$
  \end{enumerate}
\noindent That is, a random variable is a member of a p-box if its cdf $F$ falls within the cdf bounds of the p-box $F(x) \in [\underline{F}(x), \overline{F}(x)]$ for all $x$, its moments are inside the interval moments of the p-box, and it belongs to a family of distribution functions (e.g. normal, uniform) considered by the p-box. Some of the constraints may be missing. For example, if the distribution family is unknown then the set is defined solely from the cdf and moment bounds. Such p-boxes are sometimes called non-parametric, since its members do not belong to any particular class of distribution. Some constraints may also be inferred from others. For example, the interval moments may be bounded from the cdf bounds, and cdf may be bounded from moment information (explored further in section \ref{section:Moments}). Figure \ref{fig:figure1} (scripts for figures found in \texttt{/examples/JuliaCon}) shows an example of a distribution function, interval and a p-box. The grey shaded region in the interval and p-box bounds an infinite collection of probability distributions.

@@

  \caption{Various parametrically constructed p-boxes with interval parameters.}
  \label{fig:figure3}
\end{figure*}
P-boxes generalise precise distributions and intervals in the following way. A distribution is a p-box with a precise cdf and moments, i.e. when
\begin{align*}
  \underline{F}(x) &= \overline{F}(x), \\ 
  \underline{\mu}  &= \overline{\mu}, \\ 
  \underline{\sigma}^2 &= \overline{\sigma}^2 .
\end{align*}
\noindent An interval $[a,b]$ is a p-box whose bounds are step functions
\begin{align*}
    \underline{F}(x) &= \epsilon_{b}(x) ,\\
    \overline{F}(x) &= \epsilon_{a}(x) ,
\end{align*}
\noindent where $\epsilon_k$ is
\begin{equation*}
   \epsilon_k(x) = \begin{cases} 0 &\text{when } x < k \\ 1 &\text{when } x \geq k. \end{cases} 
\end{equation*}
\noindent Moreover, theoretical bounds on the mean and variance of an interval can be found \cite{ferson2002ramas}
\begin{align*}
  \mu &\in [a, b], \\
  \sigma^2 &\in [0, (b - a)^{2}/4].
\end{align*}
\noindent That is, it is not possible to find a distribution inside $[a, b]$ with a variance greater than $(b-a)^{2}/4$. The lower bound on the variance is zero, since scalars are also included in the interval. Under this definition of an interval, a random sample will always be the interval $[a, b]$, the cdf returns 0 when $x < a$, the interval $[0,1]$ when $a \leq x < b$, and $1$ when $b \leq x$. Further, the probability measure of the interval $X = [a, b]$ will be: 
\begin{equation*}
  \mathbb{P}_{X}(U) = \begin{cases}
    0 & \text{when } U \cap [a,b] = \emptyset \\
    1 & \text{when } [a,b] \subseteq U  \\
    [0, 1] & \text{otherwise.}\\
  \end{cases}
\end{equation*}
\noindent That is, if the set $U$ does not intersect the interval $X$, $\mathbb{P}_{X}=0$, i.e. $U$ certainly does not contain any of the random variables. If $X$ is fully contained in $U$, $\mathbb{P}_{X}=1$, i.e. $U$ certainly contains all of the random variables. And finally if $U$ intersects (but does not contain) $X$, the probability measure is the vacuous probability interval $\mathbb{P}_{X} = [0, 1]$, i.e. we are completely uncertain about the containment. Note that the $P_{X} = \{0, 1, [0,1]\}$ is due to the interval bounds being step functions. Generally p-boxes may yield other probability intervals.
\subsection{Outer representations of p-boxes}
\label{sec:outer_approx}
An important feature of probability bounds analysis is how distribution functions and p-boxes are represented. Generally, analytical solutions for cdfs are not readily available, for example the normal distribution's cdf can only be found by integrating the probability density, usually with quadrature. However, even if the functions were available analytically, when the variables are used in an arithmetic operation the output distribution will not necessarily belong to the same family. Therefore, we require a representation of these continuous functions which is not dependent on the distribution family (for example as used in polynomial chaos expansion), and ideally is robust in the sense that it can produce an interval error for the numerical representation. Such a representation of a probability is sometimes called an "outer approximation", in the same fashion an interval is an outer approximation of a floating point number. Such outer approximations contract to the exact value or function as more computational resources are used.

@@

A minimal p-box structure (supplementary code found in \texttt{/examples/minimal}) is as follows
\begin{lstlisting}[language = Julia]
  using IntervalArithmetic

@@

    return pbox(u_z_, d_z_, Z_moms.mean, 
    Z_moms.var, F.shape)
  end
\end{lstlisting}
\noindent How moment transformations are performed are not described here, however they are available in \texttt{MomentArithmetic.jl} \cite{ferson2021distribution}. Figure \ref{fig:figure4} shows an example of an operation between a precise distribution and an interval.
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_dist.pdf}
  \raisebox{8.0mm}{\noindent\Large*}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_in.pdf}
  \raisebox{9.0mm}{{\Large$\rightarrow$}}
  \includegraphics[width=0.12\textwidth]{../examples/JuliaCon/fig4/fig4_pbox.pdf}
  \caption{Illustration of a p-box resulting from a binary operation between a distribution and an interval.}
  \label{fig:figure4}
\end{figure}
\subsubsection{When dependencies between distributions are unknown}
In the case where a binary operation is performed between two distributions but whose dependence or correlation is unknown, a single output distribution cannot be defined. However a bounding p-box can be computed. This is the original question of Kolmogorov as described in the introduction. We call such operations between p-boxes where the marginals are fixed but no joint information is known \textit{Fréchet} operations, after the mathematician Maurice R. Fréchet, who first discovered how to bound joint distributions with fixed marginals \cite{frechet1935generalisation}. Figure \ref{fig:figure5} illustrates a Fréchet operation between a uniform and a normal distribution.
\begin{figure}[htp]

@@

A bivariate copula (2-copula) $C$ is any function $C:[0,1]^2 \rightarrow [0,  1]$ with the following properties
\begin{enumerate}
  \item Grounded: $C(0,v) = C(u,0) = 0$,
  \item Uniform margins: $C(u,1) = u; \;C(1,v) = v$,
  \item 2-increasing: \\ $C(u_{2},v_{2}) - C_{}(u_{2}, v_{1}) - C(u_{1}, v_{2}) + C(u_{1}, v_{1}) \ge 0$\\ for all $0 \le u_{1} \le u_{2} \le 1$ and $0 \le v_{1} \le v_{2} \le 1$ .
\end{enumerate}
\noindent Three important 2-copulas are
\begin{align*}
  W(u,v) &= \mathrm{max}( u + v-1,0),  \\
  \Pi(u,v) &= uv, \\
  M(u,v) &= \mathrm{min}(u,v),
\end{align*}
\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.45\textwidth]{../examples/JuliaCon/fig7/biv_cdf.pdf}\hfill
  \includegraphics[width=0.45\textwidth]{../examples/JuliaCon/fig7/samples.pdf}
  \caption{Cdf bounds of a bivariate p-box (right) with $X \sim \text{Beta}([4,8], 3)$, $Y \sim \text{N}([0,2], 2)$, and $\rho_{XY} = -0.8$, and 100 samples (left).}
  \label{fig:figure7}
\end{figure*}
\noindent where $W$ encodes maximally negative correlation, $\Pi$ encodes stochastic independence, and $M$ encodes maximally positive correlation. Moreover, $W$ and $M$ are bounds on all 2-copulas: $W(u,v) \leq C(u,v) \leq M(u,v)$. A bivariate cdf $F_{XY}$ can be constructed from two marginal cdfs $F_{X}$ and $F_{Y}$ and copula $C$ using Sklar's theorem
\begin{equation*}
  F_{XY}(x,y) = C(F_{X}(x), F_{Y}(y)),
\end{equation*}
\noindent where the constructed bivariate distribution will follow the dependence encoded in the copula and have the prescribed marginals. Sklar's theorem also works in reverse, where the dependence can be extracted from a multivariate distribution: 
\begin{equation*}
  C(u,v) = F_{XY}(F^{-1}_{X}(u), F^{-1}_{Y}(v)).
\end{equation*}
\noindent Copulas also give a very convenient way to sample from $F_{XY}$ in a generalisation of inverse transform sampling, where first the copula is sampled $(u,v) \sim C$, and then transformed through the inverses of the marginals $(F^{-1}_{X}(u), F^{-1}_{Y}(v))$.

@@

  \caption{Top left shows an initial p-box, and others show results of unary operations on this p-box.}
  \label{fig:figure8}

\end{figure*}
A minimal structure, cdf, random sample and probability measure of a bivariate p-box is
\begin{lstlisting}[language = Julia]
  using BivariateCopulas
  using BivariateCopulas: sample

@@

  \caption{Top left shows two initial p-boxes $X, Y \sim U([1,2], 4)$, and others show results of binary operations on these two p-boxes with different dependencies. Each colour represents a different copula. The colour code is $C = \{$ {\color{purple} $W$} (opposite), {\color{green} $\Phi_{-0.5}$ }(correlation -0.5), {\color{red} $\Pi$} (independence), {\color{orange} $\Phi_{0.5}$} (correlation 0.5), {\color{blue} $M$}(perfect) $\}$.}
  \label{fig:figure9}
\end{figure*}
A binary operation between two random variables $X$ and $Y$ with cdfs $F_{X}$ and $F_{Y}$ and copula $C$ can be computed with the following Lebesgue-Stieltjes integral
\begin{equation*}
  \sigma_{C,L}(z) = \int_{L\{z\}}dC(F_{X}(x),F_{Y}(y)) ,
\end{equation*}
where the integration domain is the set $L\{z\} = \{(x,y)| x,y \in \mathbb{R}, L(x,y) < z\}$ (the set of all $x$ and $y$ for which $L(x,y) < z$), and where $L$ is some binary operation (e.g. $L(x,y) = x+y$). The above $\sigma_{C,L}$ convolution is written for precise distributions, however can be extended for p-boxes by performing the convolution with some combination of the p-boxes bounds in the style of interval arithmetic, e.g.,
\begin{eqnarray*}
  + 
  \begin{cases}
    \underline{F_Z} = \underline{F_X} + \underline{F_Y} \vspace{2mm}\\
    \overline{F_Z} = \overline{F_X} + \overline{F_Y}
  \end{cases},\;
  -\begin{cases}
    \underline{F_Z} = \underline{F_X} - \overline{F_Y}\vspace{2mm}\\
    \overline{F_Z} = \overline{F_X} - \underline{F_Y}
  \end{cases} .
\end{eqnarray*}
Williamson and Downs provide an algorithm to compute rigorous bounds on the above convolution, but only under independence (when $C = \Pi$). Around the same time, Yager \cite{yager1986arithmetic} described a similar style convolution on two belief functions under the assumption independence, which is very similar to that of Williamson and Downs'. We therefore provide a generalised version of Yager's algorithm where any copula $C$ can be specified. Similar to a unary operation, a binary operation between two p-boxes can be performed by first transforming the bivariate p-box into a two dimensional belief function (which has interval boxes as focal elements), evaluating the operation on the focal elements with interval arithmetic, and transforming the resulting belief function (which will be univariate after the operation) back to a p-box. Converting a bivariate p-box into a 2D belief function is similar to that of the 1D case, however this time the mass assignment $m$ will be determined by the copula $C$. For two univariate focal elements $X$ and $Y$ with probability widths $p_{x} = [p_{x1}, p_{x2}]$ and $p_{y} = [p_{y1}, p_{y2}]$ (i.e. their vertical width in Figure \ref{fig:figure2}, with their 1D mass assignment being $p_{x2} - p_{x1}$), the 2D mass assignment of the interval box $X \times Y$ will be
\begin{equation*}
  C(p_{x2}, p_{y2}) - C(p_{x1}, p_{y2}) - C(p_{x2}, p_{y1}) + C(p_{x1}, p_{y1}) .
\end{equation*}
The following function converts a bivariate p-box into a 2D belief function
\begin{lstlisting}[language = Julia]
  function split(F :: BivPbox)
    X = F.X; Y = F.Y; C = F.C
